Towards Transparency and Open Science

### 2023-02-17 manuscript.Rmd
second batch of julia's suggestions

<!--the appraisal of "induced theories" should be low-->
So what is to be thought
Researchers and philosophers of science differ considerably in their opinion about how to appraise theories, e.g., judging the long-term performance (if they are frequentists), degrees of belief (if they are Bayesians), or probativeness [if they are severe testers, @mayoStatisticalInferenceSevere2018, p. 14] of a hypothesis.
Whatever measure they subscribe to, they would agree on a low appraisal of an untested theory.

<!--transparency about induction is required-->
So empirical researchers find themselves in a predicament.
On the one hand, they need induction to test their imprecise theories.
On the other hand, induction may render any test of a hypothesis ineffective.
Therefore, I argue that the problem is not induction but making transparent where and to what extent induction is used in the inferential process.
The replication crisis can be traced to a misjudgment of how much induction has been going on in psychology and hence, how well-tested the empirical claims, as reported in the literature, actually are.
Therefore, the question of this dissertation is *what* must be made transparent, and *how* to best make it transparent?

<!--forshadow the main part-->
The first question (the *what*) is theoretical in nature.
It is addressed in the synopsis, which supplies the theoretical framework of the articles written as part of the dissertation.
Under this framework, induction is split into a process that can be formally analyzed (statistical methods) and a part that is much more difficult to judge in current research practice (sociological factors).

<!--how it relates to the articles-->
Based on this distinction, my research articles that are part of this dissertation answer the second question (the *how*).
I argue that transparency in statistical methods is enabled by computational reproducibility, while transparency about sociological factors is facilitated by preregistration.
The conceptualization of computational reproducibility and preregistration in regard of transparency is supplemented by practical guidance on how researchers may implement these approaches.

@@

The need for transparency is closely tied to the use of induction in the empirical test of a theory.
There has been a long and vigorous debate about what it means to test a theory empirically [@popperLogicScientificDiscovery2002, Ch.\ 5 Experience as a Method].
I do not attempt to rehash the debate about what constitutes an empirical test.
However, I aim to examine the role of transparency in two frameworks that lend themselves to investigate the sources of inductive bias.
The first framework motivates transparency when an empirical test aims to evaluate a theory's verisimilitude ("closeness to truth").
The second framework motivates transparency under a science that wants to select a theory according to its expected predictive performance.

Both frameworks show how unaccounted induction leads to overconfidence in empirical results and imply some theoretical tools on how to disclose induction so that this overconfidence can be assessed and controlled.
Because both frameworks are quite technical, they are followed by a more conceptual summary of these ideas.
These sections provide the basis for understanding how computational reproducibility and preregistration enable a proper assessment of an empirical test on a conceptual level.

@@

Information theory provides a rigorous mathematical measure that can be understood as the verisimilitude of a theory [@niiniluotoVerisimilitudeThirdPeriod1998].
The distance to the truth can be formalized in terms of how much information about the truth is lost when the theory is used to model reality [@rosenkrantzMeasuringTruthlikeness1980].
Expressed mathematically, assume the existence of a function $f(x)$ that yields the likelihood of observing the state of the world $x$ where $f$ represents the ground truth.
The quantity of interst is how much information is lost if we use $g(x)$, our theory as description of the world..., instead of $f(x)$, the reality, over all possible states $\mathcal{X}$.
Expressed as lost bits of information, a measure known as Kullback–Leibler divergence [@kullbackInformationSufficiency1951], we get:
<!--x is a single state, X is the random variable, and \mathcal{X} is all possible outcomes of X-->
\begin{align} 
\mathcal{L}_{KL}(f, g) &=\int_{\mathcal{X}}f(x)\log\left(f(x)\right)\mathrm{d}x - \int_{\mathcal{X}}f(x)\log\left(g(x)\right)\mathrm{d}x\\
&= \mathbb{E}_{X \sim f}\left[\log(f(x))\right] - \mathbb{E}_{X \sim f}\left[\log(g(x))\right]\label{eq:klexp}
\end{align}

Many readers will recognize that this information-theoretic setup and the derivation below follow closely @burnhamModelSelectionMultimodel2002, Ch. 7.2, in their derivation of the Akaike Information Criterion [@hirotuguakaikeInformationTheoryExtension1971] in its general form.
What is of interest here is not the derivation but how this conceptualization may help us to understand what happens when data is simultaneously used to induce quantities of a theory and to test the theory.

@@

To allow quantities to be induced, we must assume that the theory is parameterized, e.g., $g(x|\theta)$.
That means the theory implies a family of possible probability distributions that may describe reality.
This parameterization captures the idea that some assumptions necessary for a theory to make testable statements are arbitrary.
Of those arbitrary assumptions, we want to find those that fit the reality with the least amount of information lost.
The best parameterization is achieved by:
\begin{equation}
\theta_* = \operatorname*{arg\,min}_\theta \mathcal{L}_{KL}(f, g(\cdot|\theta))
\end{equation}

@@

The crucial point is to understand what happens when $\theta$ cannot be derive deductively but must be substituted inductively with an estimate $\hat\theta(y)$.
Any estimated parameters $\hat{\theta}(y)$ would almost surely not be equal to $\theta_*$ (assuming $\theta$ may take an infinite number of values, i.e., is continues).
It follows, almost surely, that information is lost:
\begin{equation}
\mathcal{L}_{KL}(f, g(\cdot|\hat{\theta}(y)) > \mathcal{L}_{KL}(f, g(\cdot|\theta_*))
\end{equation}
Or ignoring the constant:
\begin{equation}
\mathbb{E}_{X \sim f}\left[\log(g(x|\theta_*))\right] >\mathbb{E}_{X \sim f}[\log(g(x|\hat{\theta}(y)))] 
\end{equation}

That is to say, any induced estimate will be sub-optimal.
The inference goal, however, is to compare the theory $g$ to reality $f$, not to evaluate the estimates of $\hat\theta$.
The point is to make a statement about the theory, not to make a statement about the data in light of the theory.
If the estimate of $\hat\theta(y)$, i.e., the inductive process, is unbiased in the sense that it converges towards $\theta_*$, we may form an expectation over the data $Y$:
\begin{equation}
\mathbb{E}_{Y \sim f^n}\mathbb{E}_{X \sim f}[\log(g(x|\hat\theta(y)))]\label{eq:dataexp}
\end{equation}

@@

Where $J$ is the Fisher information matrix with regard to $g$, and $I$ for $f$, respectively.
For more details about this derivation, see @burnhamModelSelectionMultimodel2002, Ch. 7.2.

The observed likelihood $\log(g(y|\hat\theta(y)))$ is, therefore, a biased estimate of the distance to the truth.
We may conclude that substituting deduced quantities with induced estimates leads to some overconfidence about how close one is to the truth.
This overconfidence is directly related to how much induction a model entails.
This bias is often called the complexity or capacity of a model, i.e., how much the data are influencing the results, hence, how detailed the model is representing the data [@goodfellowCapacityOverfittingUnderfitting2016; @mikkelsonComplexityVerisimilitudeRealism2001].
I therefore denote it as as $\mathcal{C}$.
\begin{equation}
\mathbb{E}_{X \sim f}\mathbb{E}_{Y \sim f^n}[log(g(x|\hat\theta(y)))] \approx \mathbb{E}_{Y \sim f^n}[\log(g(y|\hat{\theta}(y)))] + \mathcal{C}
\end{equation}

If we want to induce quantities and correctly appraise a theory on the same data, we must know how much we have to correct our appraisal for how the data influences the theory.
Fortunately, it is possible to approximate the complexity of a model under some conditions.
One condition is, since $\theta_*$ is unknown, that we know the properties of the inductive process that generated $\hat\theta$.
We then can formally analyze the behavior and derive a mathematical expression for $\mathcal{C}$, e.g., corrections for a large class of statistical models, most famously the class of linear models, are well known, e.g., adjusted R² [@olkinUnbiasedEstimationCertain1958], Stein's Unbiased Risk Estimator [@steinInadmissibilityUsualEstimator1956; @steinEstimationMeanMultivariate1981], Mallow's Cp [@olkinUnbiasedEstimationCertain1958; @boisbunonAkaikeInformationCriterion2014] and information criteria [@gelmanUnderstandingPredictiveInformation2014; @konoshiGeneralisedInformationCriteria1996].

## A future performance perspective

In addition to closeness to truth, there is another line of argumentation about why transparency about the process of induction is important.
Instead of verisimilitude, one might be concerned with future performance [@yarkoniChoosingPredictionExplanation2017].
That is, how well does a theory do in predicting novel facts?
Please note that the information-theoretic setup above has not appealed to the expected performance on unseen data.
Verisimilitude and expected performance are different motivations for transparency, though they can be linked.
In the future performance setup, we do not appeal to ground truth (I drop $\sim{}f$, though the samples still follow some distribution), replace the Kullback–Leibler divergence with an arbitrary loss function, and no longer require $g(x)$ to return a likelihood ($L$ stands for an arbitrary loss function):
\begin{equation}
\mathcal{L}(x, g(x|\theta)) = \mathbb{E}_xL(x, g(x|\theta))
\end{equation}

Again the loss observed in the sample used to estimate $\hat\theta$ [@sochLawUnconsciousStatistician2020] can be defined:
\begin{equation}
\mathbb{E}_y L(y, g(y|\hat\theta(y)))
 = \frac{1}{n}\sum_{i=1}^n L(y_i, g(y_i|\hat\theta(y)))
\end{equation}

However, what we are interested in is not how well the theory did on data that informed it but on future, yet unseen, data:
\begin{equation}
\mathbb{E}_x\mathbb{E}_yL(x, g(x|\hat\theta(y)))
\end{equation}
This expectation over, what is often called training and test data, is termed generalization error or expected prediction error [@bengioNoUnbiasedEstimator2004].
Note that both training ($y$) and test ($x$) data vary in this expectation.
Therefore, it is closely related to the expectation over data shown in Eq.\ \@ref(eq:dataexp) in the information-theoretic setup [@stoneAsymptoticEquivalenceChoice1977].

Instead of using the Taylor series expansion, we can repeatedly sample data and repeat the inductive process.
That is, we use cross-validation where the data is partitioned in $n$ subsets of size $n-1$ and the inductive process is repeated on each subset.
For each subset, the resulting model is then compared to the complement that was not used for induction, which is indicated by $y_{-i} = y \diagdown \{y_i\}$.
\begin{equation}
\mathbb{E}_{x}\mathbb{E}_{y}  L(x, g(x|\hat\theta(y)))
 = \frac{1}{n}\sum_{i=1}^{n}L(y_i, g(y_i|\hat\theta(y_{-i})))
\end{equation}

As stated earlier, using cross-validation, it is possible to estimate \linebreak[4] $\mathbb{E}_{X \sim f}\mathbb{E}_{Y \sim f^n}[log(g(x|\hat\theta(y)))]$ as well; which connects the information-theoretic setup with this approach [@stoneAsymptoticEquivalenceChoice1977; @stoneCrossValidatoryChoiceAssessment1974].
To make the link to the first approach even more clear, we may give an alternative definition of complexity as the expected difference between observed prediction error and expected prediction error [@hauensteinComputingAICBlackbox2018]:
\begin{equation}
\mathcal{C} = \frac{1}{n} \sum_{i=1}^{n}L(y_i, g(y_i|\hat\theta(y_{-i}))) - L(y_i, g(y_i|\hat\theta(y)))
\end{equation}

@@

Now that we have established the need for transparency about the inductive process, we can drop a few more technical details to clarify what we have to make transparent.
It bears repeating that simply revealing what has been done is not enough.
Merely showing the inductive results instead of the process that generated them is insufficient to appraise the theory.
On a conceptual level, we want to compare the following:
\begin{equation}
\mathcal{L}(\text{Theory}, \text{Reality})
\end{equation}

Where $\mathcal{L}$ stands for the loss function, i.e., how to compare predictions and reality.
To allow for induction, we replace theory with a model (not necessarily a statistical one) or, put differently, a multitude of implications about the data from the theory.

@@

So the observed loss is an underestimation that has to be corrected for by considering the complexity of the model.
\begin{equation}
\mathcal{L}(\text{Model(Reality)}, \text{Reality}) \approx \mathcal{L}(\text{Model(Sample)}, \text{Sample}) + \mathcal{C}(\text{Model})
\end{equation}

Transparency is necessary because induction leaves researchers overly optimistic regarding their theories' fit to the data ($\mathcal{C}$).
The extent of this optimism depends on the inductive process, not merely its results.
Specifically, it depends on the complexity, i.e., the ability of the inductive process to adapt to data.
The ability to adapt to data is independent of the specific data that was observed.
Complexity is a function of the model, i.e., $\mathcal{C}(\text{Model})$, not of the data.
Without knowing the inductive process, researchers cannot judge the overconfidence, so the inductive process ought to be made transparent.

@@

The above sections aimed to motivate the observation that the apparent fit of a theory to data is often overly optimistic if it has inductive elements.
This observation is only useful if we know the extent of this optimism.
However, both setups show that the extent of the optimism is closely related to the complexity of the inductive process ($\mathcal{C}(\text{Model})$) and suggest two starting points for making this bias transparent.
The first requires a formal analysis of the statistical model and the inductive process it was embedded in, to compute the complexity (parametric, e.g., using information criteria), and the second merely requires that the process is repeatable (nonparametric, e.g., using cross-validation).
Both approaches require researchers to make the inductive process transparent instead of merely publishing the results.

@@

As a first step, it can be noted that one may split $\mathcal{C}$ in complexity that can be formally described and a part that can only be evaluated subjectively.
Formal, in the strict sense, means that the complexity can be mathematically derived and in a looser sense that the process can be repeated at will.
If a researcher employs a linear model, the complexity is calculable parametrically.
Even if it were not, the linear model could be fitted on a large set of other data to asses its expected inductive behavior nonparametrically.
However, suppose the researcher reconsiders the model based on the results and their internal cognitive model (including prior knowledge, expectations, cognitive biases).
For example, based on worse-than-expected results and their intuition, they decided to add a predictor to the model.
In that case, the complexity can not be formally judged because the researcher is an intractable part of the inductive process.
The researcher cannot be asked to repeat the process on all possible data sets, nor is the behavior mathematically well-defined.
However, it is known that the resulting linear model is more complex than a linear regression formally implies.

@@

The imperative to enable proper judgment of theories is simple: induce only what is necessary and what is induced should, if possible, be done formally.
Otherwise, the supposedly objective test of the theory using hard data must be judged more subjectively than necessary.

To summarise, I separate complexity $\mathcal{C}$ according to the transparency that can theoretically be achieved.
Formal inductive processes allow full transparency in the sense that $\mathcal{C}$ can be objectively quantified.
Informal inductive processes allow only limited transparency and must be subjectively judged.

Therefore, there is a theoretical bound that limits transparency about the inductive process.
Some things cannot be made transparent in principle because some induction happens informally, so the complexity cannot be estimated with certainty.
However, there is another bound that restricts transparency further.
That is how well the inductive process, be it formal or informal, is communicated to the intended recipient.
To achieve transparency about the inductive process and ultimately about the empirical support of a theory, both bounds have to be increased.

In the following, I propose preregistration to move induction into the formal domain (pushing the first bound) and computational reproducibility to make formal induction transparent (pushing the second bound).
I argue that both practices should ultimately provide sufficient transparency to judge empirical support for theories, though they may have secondary benefits for science.

@@

Assume that a researcher only engaged in formal induction while testing a theory empirically.
If other researchers want to judge this empirical test properly, the inductive process must be communicated to them.
In other words, more than having formal induction is needed, the induction still must be made transparent in practice.
Such transparency can be achieved by computational reproducibility.
Computational reproducibility is usually defined as the ability to recreate the same results from the same data set [@claerboutElectronicDocumentsGive1992; @peikertReproducibleDataAnalysis2021; @peikertReproducibleResearchTutorial2021].
However, through the lens of transparency about the inductive process, it may be possible to further refine our view on computational reproducibility.

Consider software that delivers the same results upon input of the same data, i.e., it is reproducible.
However, place two restrictions on the software.
First, while able to compute results based on data, the software cannot be understood by human researchers, e.g., because it is an executable program, of which the source code may not be openly available.
Second, when the data set changes, the software does not work anymore.
The first condition rules out that the software can be analyzed formally.
The second condition prevents from changing the data and observing the results.
Therefore, there are no means to assess the complexity of the inductive process and the empirical support cannot be judged.
It follows that to provide transparency about the complexity of an inductive process, the definition of computational reproducibility must be broadened, i.e., it has to be ruled out one of the above conditions.

Suppose the second condition is relaxed; then there is a "black box" that is repeatable on other data.
In that case, there still is transparency about the inductive process because the complexity of the inductive process can be estimated, i.e., using cross-validation.
Relaxing the first condition, missing transparency for human researchers, most likely leads to relaxing the second condition, repeatability on other data, as well.
If humans can understand the inductive process, complexity can either be calculated analytically or the process can be reimplemented to work on other similar data.
Therefore, the traditional definition of computational reproducibility has to be extended and further it is required that the process can be repeated on other similar data.

@@

These two requirements are not easy to meet in practice.
In @peikertReproducibleDataAnalysis2021, I proposed a workflow that unifies both requirements under the objective of automating the full process from data to manuscript.
Implementing computational reproducibility by automating the process from data to results fulfills both conditions needed for transparency while reducing the effort that must be invested.
Verifying that the results are actually produced by the inductive process and data is then a task that a computer may fulfil without human intervention.
Removing the human from the loop facilitates that the data can be easily substituted by similar data.
Given enough computing power, automatic reproducibility allows to scale the inductive process to many data sets and therefore enables the assesment of the complexity.

While this workflow provides transparency and facilitates that the recipients of the work can judge the inductive process easily, it asks a lot from the researchers who create the work.
To address this problem and simplify the workflow's application in practice, I developed the R package `repro` [@R-repro].
I further refined and simplified the workflow to make it more accessible in @peikertReproducibleResearchTutorial2021 and contributed to @vanlissaWORCSWorkflowOpen2021's manuscript and the there presented software `worcs`.

>>>>

<!--the appraisal of "induced theories" should be low-->
So what is to be thought of such as yet untested theory?
Researchers and philosophers of science differ considerably in their opinion about how to appraise theories, e.g., judging the long-term performance (if they are frequentists), degrees of belief (if they are Bayesians), or probativeness [if they are severe testers, @mayoStatisticalInferenceSevere2018, p. 14] of a hypothesis.
Whatever measure they subscribe to, they would agree on a low appraisal of an untested theory.

<!--transparency about induction is required-->
Empirical researchers thus find themselves in a predicament.
On the one hand, they need induction to test their imprecise theories.
On the other hand, induction may render any test of a hypothesis ineffective.
Therefore, I argue that the problem is not induction but making transparent where and to what extent induction is used in the inferential process.
The replication crisis can be traced to a misjudgment of how much induction has been going on in psychology and hence, how well-tested the empirical claims, as reported in the literature, actually are.
Therefore, the questions of this dissertation are *what* must be made transparent, and *how* to best make it transparent?

<!--forshadow the main part-->
The first question (the *what*) is theoretical in nature.
It is addressed in this synopsis, which supplies the theoretical framework of the articles written as part of the dissertation.
Under this framework, induction is split into a process that can be formally analyzed (statistical methods) and a part that is much more difficult to judge in current research practice (sociological factors).

<!--how it relates to the articles-->
Based on this distinction, my research articles that are part of this dissertation answer the second question (the *how*).
I argue that transparency in statistical methods is enabled by computational reproducibility, while transparency about sociological factors is facilitated by preregistration.
The conceptualization of computational reproducibility and preregistration from a transparency perspective is supplemented by practical guidance on how researchers can implement these approaches.

@@

The need for transparency is closely tied to the use of induction in the empirical test of a theory.
There has been a long and vigorous debate about what it means to test a theory empirically [@popperLogicScientificDiscovery2002, Ch.\ 5, Experience as a Method].
I do not attempt to rehash the debate about what constitutes an empirical test.
However, I aim to examine the role of transparency in two frameworks that lend themselves to investigate the sources of inductive bias.
The first framework motivates transparency when an empirical test aims to evaluate a theory's verisimilitude ("closeness to truth").
The second framework motivates transparency under a science that wants to select a theory according to its expected predictive performance.

Both frameworks show how unaccounted induction leads to overconfidence in empirical results and imply some theoretical tools to disclose induction so that this overconfidence can be assessed and controlled.
Because both frameworks are quite technical, they are followed by a more conceptual summary of these ideas.
These sections provide the basis for understanding how computational reproducibility and preregistration enable a proper assessment of an empirical test on a conceptual level.

@@

Information theory provides a rigorous mathematical measure that can be understood as the verisimilitude of a theory [@niiniluotoVerisimilitudeThirdPeriod1998].
The distance to the truth can be formalized in terms of how much information about the truth is lost when the theory is used to model reality [@rosenkrantzMeasuringTruthlikeness1980].
Expressed mathematically, assume the existence of a function $f(x)$ that yields the likelihood of observing the state of the world $x$ where $f$ represents the ground truth.
The quantity of interest is how much information is lost if we use $g(x)$, our theory as description of the world, instead of $f(x)$, the reality, over all possible states $\mathcal{X}$.
Expressed as lost bits of information, a measure known as Kullback–Leibler divergence [@kullbackInformationSufficiency1951], we get:
<!--x is a single state, X is the random variable, and \mathcal{X} is all possible outcomes of X-->
\begin{align} 
\mathcal{L}_{KL}(f, g) &=\int_{\mathcal{X}}f(x)\log\left(f(x)\right)\mathrm{d}x - \int_{\mathcal{X}}f(x)\log\left(g(x)\right)\mathrm{d}x\\
&= \mathbb{E}_{X \sim f}\left[\log(f(x))\right] - \mathbb{E}_{X \sim f}\left[\log(g(x))\right]\label{eq:klexp}
\end{align}

Many readers will recognize that this information-theoretic setup and the derivation below closely follow @burnhamModelSelectionMultimodel2002, Chapter 7.2, in their derivation of the Akaike Information Criterion [@hirotuguakaikeInformationTheoryExtension1971] in its general form.
What is of interest here is not the derivation but how this conceptualization may help us to understand what happens when data is simultaneously used to induce quantities of a theory and to test the theory.

@@

To allow quantities to be induced, we must assume that the theory is parameterized, e.g., $g(x|\theta)$.
That means that the theory implies a family of possible probability distributions that may describe reality.
This parameterization captures the idea that some assumptions necessary for a theory to make testable statements are arbitrary.
Of those arbitrary assumptions, we want to find those that fit the reality with the least amount of information lost.
The best parameterization is achieved by:
\begin{equation}
\theta_* = \operatorname*{arg\,min}_\theta \mathcal{L}_{KL}(f, g(\cdot|\theta))
\end{equation}

@@

The crucial point is to understand what happens when $\theta$ cannot be derived deductively but must be substituted inductively with an estimate $\hat\theta(y)$.
Any estimated parameters $\hat{\theta}(y)$ would almost surely not be equal to $\theta_*$ (assuming $\theta$ may take an infinite number of values, i.e., is continuous).
It follows, almost surely, that information is lost:
\begin{equation}
\mathcal{L}_{KL}(f, g(\cdot|\hat{\theta}(y)) > \mathcal{L}_{KL}(f, g(\cdot|\theta_*))
\end{equation}
Or ignoring the constant:
\begin{equation}
\mathbb{E}_{X \sim f}\left[\log(g(x|\theta_*))\right] >\mathbb{E}_{X \sim f}[\log(g(x|\hat{\theta}(y)))] 
\end{equation}

That is to say, any induced estimate will be suboptimal.
The inference goal, however, is to compare the theory $g$ to reality $f$, not to evaluate the estimates of $\hat\theta$.
The point is to make a statement about the theory, not to make a statement about the data in light of the theory.
If the estimate of $\hat\theta(y)$, i.e., the inductive process, is unbiased in the sense that it converges towards $\theta_*$, we may form an expectation over the data $Y$:
\begin{equation}
\mathbb{E}_{Y \sim f^n}\mathbb{E}_{X \sim f}[\log(g(x|\hat\theta(y)))]\label{eq:dataexp}
\end{equation}

@@

Where $J$ is the Fisher information matrix with regard to $g$, and $I$ for $f$, respectively.
For more details about this derivation, see @burnhamModelSelectionMultimodel2002, Chapter 7.2.

The observed likelihood $\log(g(y|\hat\theta(y)))$ is, therefore, a biased estimate of the distance to the truth.
We may conclude that substituting deduced quantities by induced estimates leads to some overconfidence about how close one is to the truth.
This overconfidence is directly related to how much induction a model entails.
This bias is often called the complexity or capacity of a model, i.e., how much the data are influencing the results, hence, in how much detail the model may represent the data [@goodfellowCapacityOverfittingUnderfitting2016, Chapter. 5.2; @mikkelsonComplexityVerisimilitudeRealism2001].
I therefore denote it as as $\mathcal{C}$.
\begin{equation}
\mathbb{E}_{X \sim f}\mathbb{E}_{Y \sim f^n}[log(g(x|\hat\theta(y)))] \approx \mathbb{E}_{Y \sim f^n}[\log(g(y|\hat{\theta}(y)))] + \mathcal{C}
\end{equation}

If we want to induce quantities and correctly appraise a theory on the same data, we must know how much we have to correct our appraisal for how the data influences the theory.
Fortunately, it is possible to approximate the complexity of a model under some conditions.
Since $\theta_*$ is unknown, one condition is, that we know the properties of the inductive process that generated $\hat\theta$.
We then can formally analyze the behavior and derive a mathematical expression for $\mathcal{C}$, e.g., corrections for a large class of statistical models, most famously the class of linear models, are well known, e.g., adjusted R² [@olkinUnbiasedEstimationCertain1958], Stein's Unbiased Risk Estimator [@steinInadmissibilityUsualEstimator1956; @steinEstimationMeanMultivariate1981], Mallow's $C_p$ [@olkinUnbiasedEstimationCertain1958; @boisbunonAkaikeInformationCriterion2014] and information criteria [@gelmanUnderstandingPredictiveInformation2014; @konoshiGeneralisedInformationCriteria1996].

## A future-performance perspective

In addition to closeness to truth, there is another line of argumentation about why transparency about the process of induction is important.
Instead of verisimilitude, one might be concerned with future performance [@yarkoniChoosingPredictionExplanation2017].
That is, how well does a theory do in predicting novel facts?
Please note that the information-theoretic setup above has not appealed to the expected performance on unseen data.
Verisimilitude and expected performance are different motivations for transparency, though they can be linked.
In the future-performance setup, we do not appeal to ground truth (I drop $\sim{}f$, though the samples still follow some distribution), replace the Kullback–Leibler divergence with an arbitrary loss function, and no longer require $g(x)$ to return a likelihood ($L$ stands for an arbitrary loss function):
\begin{equation}
\mathcal{L}(x, g(x|\theta)) = \mathbb{E}_xL(x, g(x|\theta))
\end{equation}

Again the loss observed in the sample used to estimate $\hat\theta$ [@sochLawUnconsciousStatistician2020, Chapter 1.5.8] can be defined:
\begin{equation}
\mathbb{E}_y L(y, g(y|\hat\theta(y)))
 = \frac{1}{n}\sum_{i=1}^n L(y_i, g(y_i|\hat\theta(y)))
\end{equation}

However, what we are interested in is not how well the theory did on data that informed it, but on future, yet unseen, data:
\begin{equation}
\mathbb{E}_x\mathbb{E}_yL(x, g(x|\hat\theta(y)))
\end{equation}
This expectation over, what is often called training and test data, is termed generalization error or expected prediction error [@bengioNoUnbiasedEstimator2004].
Note that both training ($y$) and test ($x$) data vary in this expectation.
Therefore, it is closely related to the expectation over data shown in Eq.\ \@ref(eq:dataexp) in the information-theoretic setup [@stoneAsymptoticEquivalenceChoice1977].

Instead of using the Taylor series expansion, we can repeatedly sample data and repeat the inductive process.
That is, we use cross-validation where the data are partitioned in $n$ subsets of size $n-1$ and the inductive process is repeated on each subset.
For each subset, the resulting model is then compared to the complement that was not used for induction, which is indicated by $y_{-i} = y \diagdown \{y_i\}$.
\begin{equation}
\mathbb{E}_{x}\mathbb{E}_{y}  L(x, g(x|\hat\theta(y)))
 = \frac{1}{n}\sum_{i=1}^{n}L(y_i, g(y_i|\hat\theta(y_{-i})))
\end{equation}

As stated earlier, using cross-validation, it is possible to estimate \linebreak[4] $\mathbb{E}_{X \sim f}\mathbb{E}_{Y \sim f^n}[log(g(x|\hat\theta(y)))]$ as well; this connects the information-theoretic setup with this approach [@stoneAsymptoticEquivalenceChoice1977; @stoneCrossValidatoryChoiceAssessment1974].
To make the link to the first approach even more clear, we may give an alternative definition of complexity as the expected difference between observed prediction error and expected prediction error [@hauensteinComputingAICBlackbox2018]:
\begin{equation}
\mathcal{C} = \frac{1}{n} \sum_{i=1}^{n}L(y_i, g(y_i|\hat\theta(y_{-i}))) - L(y_i, g(y_i|\hat\theta(y)))
\end{equation}

@@

Now that we have established the need for transparency about the inductive process, we can glance over of the more technical details to clarify what we have to make transparent.
It bears repeating that simply revealing what has been done is not enough.
Merely showing the inductive results instead of the process that generated them is insufficient to appraise the theory.
On a conceptual level, we want to compare the following:
\begin{equation}
\mathcal{L}(\text{Theory}, \text{Reality})
\end{equation}

where $\mathcal{L}$ stands for the loss function, i.e., how to compare predictions and reality.
To allow for induction, we replace theory with a model (not necessarily a statistical one) or, put differently, a multitude of implications about the data from the theory.

@@

The observed loss is thus an underestimation that has to be corrected for by considering the complexity of the model.
\begin{equation}
\mathcal{L}(\text{Model(Reality)}, \text{Reality}) \approx \mathcal{L}(\text{Model(Sample)}, \text{Sample}) + \mathcal{C}(\text{Model})
\end{equation}

Transparency is necessary because induction leaves researchers overly optimistic regarding their theories' fit to the data ($\mathcal{C}$).
The extent of this optimism depends on the inductive process, not merely its results.
Specifically, it depends on the complexity, i.e., the ability of the inductive process to adapt to data.
The ability to adapt to data is independent of the specific data that were observed.
Complexity is a function of the model, i.e., $\mathcal{C}(\text{Model})$, not of the data.
Without knowing the inductive process, researchers cannot judge the overconfidence, so the inductive process ought to be made transparent.

@@

The above sections aimed to motivate the observation that the apparent fit of a theory to data is often overly optimistic if it has inductive elements.
This observation is only useful if we know the extent of this optimism.
However, both setups show that the extent of the optimism is closely related to the complexity of the inductive process ($\mathcal{C}(\text{Model})$) and suggest two starting points for making this bias transparent.
The first requires a formal analysis of the statistical model and the inductive process it was embedded in, to compute the complexity (parametric, e.g., using information criteria), and the second merely requires that the process is repeatable (nonparametric, e.g., using cross-validation).
Both approaches require researchers to make the inductive process transparent rather than merely publishing the results.

@@

As a first step, it can be noted that one may split $\mathcal{C}$ in complexity that can be formally described and a part that can only be evaluated subjectively.
Formal discription, in the strict sense, means that the complexity can be mathematically derived, and in a looser sense, that the process can be repeated at will.
If a researcher employs a linear model, the complexity is calculable parametrically.
Even if it were not calculable, the linear model could be fitted on a large set of other data to assess its expected inductive behavior nonparametrically.
However, suppose the researcher reconsiders the model based on their results and their internal cognitive model (including prior knowledge, expectations, cognitive biases).
For example, based on worse-than-expected results and their intuition, they decided to add a predictor to the model.
In that case, the complexity can not be formally judged because the researcher is an intractable part of the inductive process.
The researcher cannot be asked to repeat the process on all possible data sets, nor is their behavior mathematically well-defined.
However, it is known that the resulting linear model is more complex than a linear regression formally implies.

@@

The imperative to enable proper judgment of theories is simple: induce only what is necessary, and what is induced should, if possible, be done formally.
Otherwise, the supposedly objective test of the theory using hard data must be judged more subjectively than necessary.

To summarize, I separate complexity $\mathcal{C}$ according to the transparency that can theoretically be achieved.
Formal inductive processes allow full transparency in the sense that $\mathcal{C}$ can be objectively quantified.
Informal inductive processes allow only limited transparency and must be subjectively judged.

Therefore, there is a theoretical bound that limits transparency about the inductive process.
Some things cannot be made transparent in principle because some induction happens informally, and the complexity cannot be estimated with certainty.
However, there is another bound that restricts transparency further.
That is how well the inductive process, be it formal or informal, is communicated to the intended recipient.
To achieve transparency about the inductive process and ultimately about the empirical support of a theory, both boundaries have to be increased.

In the following, I propose preregistration as a means to move induction into the formal domain (pushing the first boundary) and computational reproducibility to make formal induction transparent (pushing the second boundary).
I argue that both practices should ultimately provide sufficient transparency to judge empirical support for theories, though they may have secondary benefits for science.

@@

Assume that a researcher only engaged in formal induction while testing a theory empirically.
If other researchers want to judge this empirical test properly, the inductive process must be communicated to them.
In other words, formal induction is needed, but it still must be made transparent in practice.
Such transparency can be achieved by computational reproducibility.
Computational reproducibility is usually defined as the ability to recreate the same results from the same data set [@claerboutElectronicDocumentsGive1992; @peikertReproducibleDataAnalysis2021; @peikertReproducibleResearchTutorial2021].
However, from the perspective of transparency about the inductive process, it may be possible to further refine our view on computational reproducibility.

Consider software that delivers the same results upon input of the same data, i.e., it enables reproducible analyses.
However, place two restrictions on the software.
First, while able to compute results based on data, the software cannot be understood by human researchers, e.g., because it is an executable program, of which the source code may not be openly available.
Second, when the data set changes, the software does not work anymore.
The first condition rules out that the software can be analyzed formally.
The second condition prevents changes of the data and observation of the results.
Therefore, there are no means to assess the complexity of the inductive process and the empirical support cannot be judged.
It follows that to provide transparency about the complexity of an inductive process, the definition of computational reproducibility must be broadened, i.e., at least one of the above conditions needs to be ruled out..

Suppose the second condition is relaxed; then there is a "black box" that is repeatable on other data.
In that case, there still is transparency about the inductive process because the complexity of the inductive process can be estimated, i.e., using cross-validation.
Relaxing the first condition, missing transparency for human researchers, most likely leads to relaxation of the second condition, repeatability on other data, as well.
If humans can understand the inductive process, complexity can either be calculated analytically or the process can be reimplemented to work on similar data.
Therefore, the traditional definition of computational reproducibility has to be extended.
Further, it is required that the process can be repeated on other similar data.

@@

These two requirements are not easy to meet in practice.
In @peikertReproducibleDataAnalysis2021, I proposed a workflow that unifies both requirements under the objective of automating the full process from data to manuscript.
Implementing computational reproducibility by automating the process from data to results fulfills both conditions needed for transparency while reducing the effort that must be invested.
Verifying that the results are actually produced by the inductive process and data is then a task that a computer may fulfill without human intervention.
Removing the human from the loop facilitates that the data can be easily substituted by similar data.
Given enough computing power, automatic reproducibility allows one to scale the inductive process to many data sets and therefore enables the assessment of the complexity $\mathcal{C}$.

While this workflow provides transparency and facilitates judgement of the inductive process by recipients of the work (readers, collaborators, editors, etc.), it asks a lot of the researchers who creating it.
To address this problem and simplify the workflow's application in practice, I developed the R package `repro` [@R-repro].
I further refined and simplified the workflow to make it more accessible in @peikertReproducibleResearchTutorial2021 and contributed to @vanlissaWORCSWorkflowOpen2021's manuscript and the software `worcs` presented there.
